### 2021-03-24 02_abstract/abstract.tex
Further improvements of text in chapter 04

In a first step, the climate sensitivity (\ie{} the temperature response of the
climate system to an external forcing) in the latest generation of \acsp{ESM}
from \acs{CMIP}6 is evaluated. For the \acfAbstract{ECS}, which is defined as
the equilibrium temperature response that follows a doubling of the atmospheric
\acfAbstract{CO2} concentration, a \acfAbstract{MMM} of $3.74 \unit{K}$ and a
multi-model range of $\rangeunit{1.8}{5.6}{K}$ are found. These values are
higher than in any previous \acs{CMIP} ensemble before. Moreover, a third of
the analyzed \acs{CMIP}6 models exceeds the upper bound of the likely \acs{ECS}
range of $\rangeunit{1.5}{4.5}{K}$ assessed by the \acfAbstract{IPCC} Fifth
\acl{AR} (\acs{AR}5) from 2013. Similarly, the transient response of the
climate system to a doubling of \acs{CO2}, also known as \acfAbstract{TCR},
shows an inter-model range of $\rangeunit{1.3}{3.0}{K}$ with an upper bound
again higher than the likely range assessed in \acs{AR}5 of
$\rangeunit{1.0}{2.5}{K}$. A possible reason for the increased climate
sensitivity in many \acs{CMIP}6 models is a change in the microphysical
representation of mixed-phase clouds over the Southern Ocean, which reduces the
strong negative shortwave cloud phase change feedback in this region that is
present in climate models from previous \acs{CMIP} generations.

>>>>

In a first step, the climate sensitivity (\ie{} the temperature response of the
climate system to an external forcing) in the latest generation of \acsp{ESM}
from \acs{CMIP}6 is evaluated. For the \acfAbstract{ECS}, which is defined as
the equilibrium temperature response that follows a doubling of the atmospheric
\acfAbstract{CO2} concentration, a \acfAbstract{MMM} of $3.74 \unit{K}$ and a
multi-model range of $\rangeunit{1.8}{5.6}{K}$ are found. These values are
higher than in any previous \acs{CMIP} ensemble before. Moreover, a third of
the analyzed \acs{CMIP}6 models exceeds the upper bound of the likely \acs{ECS}
range of $\rangeunit{1.5}{4.5}{K}$ assessed by the \acfAbstract{IPCC} Fifth
\acl{AR} (\acs{AR}5) from 2013. Similarly, the transient response of the
climate system to a doubling of \acs{CO2}, also known as \acfAbstract{TCR},
shows an inter-model range of $\rangeunit{1.3}{3.0}{K}$ with an upper bound
again higher than the likely range assessed in \acs{AR}5 of
$\rangeunit{1.0}{2.5}{K}$. Possible reasons for the increased climate
sensitivity in many \acs{CMIP}6 models are the addition of new prognostic
aerosol schemes that include aerosol-cloud interactions and changes in the
microphysical representation of mixed-phase clouds over the Southern Ocean,
which reduce the strong negative shortwave cloud phase change feedback in this
region that is present in climate models from previous \acs{CMIP} generations.

### 2021-03-24 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Further improvements of text in chapter 04

Various improvements of the underlying physical, biological and chemical
processes have been introduced to the \acs{CMIP}6 \acp{ESM} in order to
represent the coupled Earth system in more detail. Since many of these
processes directly influence the models' \ac{ECS} \autocite{Forster2020},
determining reasons for the increased \ac{ECS} in the \acs{CMIP}6 ensemble is
highly non-trivial. As discussed in detail by \textcite{Meehl2020}, one
possible reason is the consideration of aerosol-cloud interactions. New
prognostic aerosol schemes added to some \acs{CMIP}6 models that include
aerosol-cloud interactions could have produced overly large negative radiative
forcing, which then required a stronger model response to \ac{GHG} forcing in
order to reproduce the observed historical temperature trend. In fact,
\textcite{Meehl2020} find a weak inter-model relationship between the aerosol
forcing and \ac{ECS} for some \acs{CMIP}6 models, with larger negative
present-day aerosol forcing associated with higher \ac{ECS}. Such relationships
have also been found for previous climate model generations
\autocite{Kiehl2007, Forster2013}. However, due to the varying aerosol forcing
over time, these relations cannot be used to tune \ac{ECS} based on the models'
responses to aerosols over the different periods of the \nth{20} century
\autocite{Dittus2020}. An example for a \acs{CMIP}6 model with high \ac{ECS}
and a new aerosol scheme is the \ac{HadGEM} version 3. The new aerosol scheme
in this model suppresses a strong negative cloud feedback over the Southern
Ocean \autocite{BodasSalcedo2019}, which is one of the most pristine aerosol
regions in the world \autocite{Hamilton2014}, where aerosols are predominantly
of natural origin (\eg{} marine emissions of gaseous dimethyl sulfide). Earlier
\ac{HadGEM} versions show a large reduction in cloud droplet size with warming
over the Southern Ocean, which produces a negative shortwave cloud feedback.
However, in \acs{HadGEM}3, this negative feedback is completely suppressed or
even positive due to a small increase of the cloud droplet size with warming,
which eventually increases \ac{ECS} due to an increased net cloud feedback
\autocite{BodasSalcedo2019}.

@@

However, it is well known that the low-level mixed-phase clouds over the
Southern Ocean in these earlier climate models are biased towards too large
amounts of ice crystals and too little amounts of supercooled liquid water when
compared to satellite observations \autocite{BodasSalcedo2016}. Thus, the size
of the resulting strong negative cloud phase change feedback in these models
has long been questioned \autocite{McCoy2015, Tan2016}. Improvements in the
representation of mixed-phase clouds are known to reduce this long-standing
cloud phase bias over the Southern Ocean \autocite{BodasSalcedo2016, McCoy2016}
and to improve the representation of both cloud microphysical structure and
cloud radiative impacts in this region \autocite{Hyder2018, Kay2016}. For this
reason, the microphysical representation of mixed-phase clouds has been updated
in many \acs{CMIP}6 models. For example, in \acs{CESM}2, a new mixed-phase ice
nucleation scheme by \textcite{Hoose2010} replaces the \textcite{Meyers1992}
empirical scheme used in predecessor models \autocite{Gettelman2019}.
Simulations with the \textcite{Meyers1992} scheme show almost no supercooled
liquid water in high-latitude clouds and a strong negative cloud phase feedback
when ice turns to liquid. On the contrary, simulations with the new ice
nucleation scheme by \textcite{Hoose2010} show a strong reduction in this
feedback and are more consistent with observations of the current climate
\autocite{Gettelman2019}. Another example is the \acs{HadGEM}3 model, for which
a new cloud scheme that accounts for turbulent production of liquid water in
mixed-phase clouds has been implemented \autocite{BodasSalcedo2019}. The new
scheme parameterizes the role of subgrid-scale turbulence in the production and
maintenance of supercooled liquid water \autocite{Furtado2016}. As a
consequence, the cloud liquid water content over the Southern ocean is
increased, which is in closer agreement to observations compared to earlier
\ac{HadGEM} versions. Moreover, the increase in the cloud liquid water content
with warming is reduced in \acs{HadGEM}3, leading to a reduction in the
negative cloud phase change feedback over the Southern Ocean
\autocite{BodasSalcedo2019}. By evaluating short-term forecasts created with
\acs{HadGEM}3, \textcite{Williams2020} additionally show that the new climate
model version improves in traditional metrics used for numerical weather
prediction, which indicates an improvement in the physical processes simulated
by the climate model.

>>>>

Various improvements of the underlying physical, biological and chemical
processes have been introduced to the \acs{CMIP}6 \acp{ESM} in order to
represent the coupled Earth system in more detail. Since many of these
processes directly influence the models' \ac{ECS} \autocite{Forster2020},
determining reasons for the increased \ac{ECS} in the \acs{CMIP}6 ensemble is
highly non-trivial. As discussed in detail by \textcite{Meehl2020}, one
possible reason is the consideration of aerosol-cloud interactions. New
prognostic aerosol schemes added to some \acs{CMIP}6 models that include
aerosol-cloud interactions could have produced overly large negative radiative
forcing, which then required a stronger model response to \ac{GHG} forcing in
order to reproduce the observed historical temperature trend. In fact,
\textcite{Meehl2020} find a weak inter-model relationship between the aerosol
forcing and \ac{ECS} for some \acs{CMIP}6 models, with larger negative
present-day aerosol forcing associated with higher \ac{ECS}. Such relationships
have also been found for previous climate model generations
\autocite{Kiehl2007, Forster2013}. However, due to the varying aerosol forcing
over time, these relations cannot be used to tune \ac{ECS} based on the models'
responses to aerosols over the different periods of the \nth{20} century
\autocite{Dittus2020}. A further type of aerosol-cloud interactions is for
example present in the \ac{HadGEM} version 3. The new aerosol scheme in this
model suppresses a strong negative cloud feedback over the Southern Ocean that
is found in the model's predecessor versions \autocite{BodasSalcedo2019}. The
Southern Ocean is one of the most pristine aerosol regions in the world
\autocite{Hamilton2014}, where aerosols predominantly originate from natural
sources like marine emissions of gaseous dimethyl sulfide. Earlier \ac{HadGEM}
versions show a large reduction in cloud droplet size with warming over the
Southern Ocean, which increases the cloud albedo and thus produces a negative
shortwave cloud feedback. On the contrary, in \acs{HadGEM}3, this negative
feedback is completely suppressed or even positive due to a small increase of
the cloud droplet size with warming, which eventually increases \ac{ECS} due to
an increased net cloud feedback \autocite{BodasSalcedo2019}.

@@

However, it is well known that the low-level mixed-phase clouds over the
Southern Ocean in these earlier climate models are biased towards too large
amounts of ice crystals and too little amounts of supercooled liquid water when
compared to satellite observations \autocite{BodasSalcedo2016}. Thus, the size
of the resulting strong negative cloud phase change feedback in these models
has long been questioned \autocite{McCoy2015, Tan2016}. Improvements in the
representation of mixed-phase clouds are known to reduce this long-standing
cloud phase bias over the Southern Ocean \autocite{BodasSalcedo2016, McCoy2016}
and to improve the representation of both cloud microphysical structure and
cloud radiative impacts in this region \autocite{Hyder2018, Kay2016}. For this
reason, the microphysical representation of mixed-phase clouds has been updated
in many \acs{CMIP}6 models. For example, in \acs{CESM}2, a new mixed-phase ice
nucleation scheme by \textcite{Hoose2010} replaces the \textcite{Meyers1992}
empirical scheme used in predecessor models \autocite{Gettelman2019}.
Simulations with the \textcite{Meyers1992} scheme show almost no supercooled
liquid water in high-latitude clouds and a strong negative cloud phase feedback
when ice turns to liquid. On the contrary, simulations with the new ice
nucleation scheme by \textcite{Hoose2010} show a strong reduction in this
feedback and are more consistent with observations of the current climate
\autocite{Gettelman2019}. Another example is the \acs{HadGEM}3 model, for which
a new cloud scheme that accounts for turbulent production of liquid water in
mixed-phase clouds has been implemented \autocite{BodasSalcedo2019}. The new
scheme parameterizes the role of subgrid-scale turbulence in the production and
maintenance of supercooled liquid water \autocite{Furtado2016}. As a
consequence, the cloud liquid water content over the Southern Ocean is
increased, which is in closer agreement to observations compared to earlier
\ac{HadGEM} versions. Moreover, the increase in the cloud liquid water content
with warming is reduced in \acs{HadGEM}3, leading to a reduction in the
negative cloud phase change feedback over the Southern Ocean
\autocite{BodasSalcedo2019}. By evaluating traditional metrics used for
numerical weather prediction, \textcite{Williams2020} additionally demonstrate
that short-term forecasts created with \acs{HadGEM}3 show a higher performance
than comparable simulations of predecessor model versions, which indicates an
improvement of the physical processes in the new climate model.

### 2021-03-24 ch07_conclusion/conclusion.tex
Further improvements of text in chapter 04

In the first study of this thesis, the climate sensitivity metrics \ac{ECS} and
\ac{TCR} are evaluated for the latest generation of \acp{ESM} from \acs{CMIP}6.
This work, which is presented in \cref{ch:04:papers_ecs_tcr_assessment} and
already published in \textcite{Bock2020, Meehl2020}, directly addresses
\cref{enum:01:question_1} (\enquote{\emph{\KeyScienceQuestionOne{}}}). For
\ac{ECS}, a \acs{CMIP}6 model range of $\rangeunit{1.8}{5.6}{K}$ is found,
which is higher on the upper and lower end than any model range from previous
\ac{CMIP} generations before. In comparison to \acs{CMIP}5, the \acs{CMIP}6
\ac{MMM} of \ac{ECS} is about $16 \unit{\%}$ higher ($3.74 \unit{K}$ in
\acs{CMIP}6 versus $3.23 \unit{K}$ in \acs{CMIP}5). Moreover, especially the
upper bound of the \acs{CMIP}6 models range is considerably larger than in the
assessed range of $\rangeunit{1.5}{4.5}{K}$ given by the latest published
\ac{AR} of the \ac{IPCC} from 2013 \autocite{Stocker2013}. The assessed upper
bound of $4.5 \unit{K}$ is exceeded by a third of the \acs{CMIP}6 models, with
many models showing \ac{ECS} values above $5 \unit{K}$. For \ac{TCR}, the model
range of \acs{CMIP}6 is $\rangeunit{1.3}{3.0}{K}$, which also exceeds the
\acs{CMIP}5 range of $\rangeunit{1.1}{2.5}{K}$ and the assessed range from
\acs{AR}5 of $\rangeunit{1.0}{2.5}{K}$. One possible reason for the increased
climate sensitivity in many \acs{CMIP}6 models is the addition of prognostic
aerosol schemes that include aerosol-cloud interactions which could have
produced overly large negative radiative forcing in some models. To correctly
reproduce the observed historical temperature trend, a stronger model response
to \ac{GHG} forcing is required, resulting in a higher climate sensitivity. A
further reason for the high climate sensitivity in \acs{CMIP}6 is a change in
the microphysical representation of mixed-phase clouds over the Southern Ocean
in some models. This change was mainly implemented to address a long-standing
cloud phase bias over the Southern Ocean in the climate models, which has been
successfully reduced. Moreover, the \acs{CMIP}6 \ac{MMM} shows an improved
simulation of the shortwave \ac{CRE} when compared to observations. However,
the change in the representation of mixed-phase clouds also substantially
reduces the strong negative shortwave cloud feedback over the Southern Ocean
that is present in previous \ac{CMIP} generations and that results from a cloud
phase change from ice clouds in the present-day to liquid clouds in the future.
In the affected \acs{CMIP}6 models, this cloud phase change due to warming is
reduced since these models simulate less cloud ice over the present-day
Southern Ocean than their predecessor versions with no supercooled cloud
liquid.

>>>>

In the first study of this thesis, the climate sensitivity metrics \ac{ECS} and
\ac{TCR} are evaluated for the latest generation of \acp{ESM} from \acs{CMIP}6.
This work, which is presented in \cref{ch:04:papers_ecs_tcr_assessment} and
already published in \textcite{Bock2020, Meehl2020}, directly addresses
\cref{enum:01:question_1} (\enquote{\emph{\KeyScienceQuestionOne{}}}). For
\ac{ECS}, a \acs{CMIP}6 model range of $\rangeunit{1.8}{5.6}{K}$ is found,
which is higher on the upper and lower end than any model range from previous
\ac{CMIP} generations before. In comparison to \acs{CMIP}5, the \acs{CMIP}6
\ac{MMM} of \ac{ECS} is about $16 \unit{\%}$ higher ($3.74 \unit{K}$ in
\acs{CMIP}6 versus $3.23 \unit{K}$ in \acs{CMIP}5). Moreover, especially the
upper bound of the \acs{CMIP}6 models range is considerably larger than in the
assessed range of $\rangeunit{1.5}{4.5}{K}$ given by the latest published
\ac{AR} of the \ac{IPCC} from 2013 \autocite{Stocker2013}. The assessed upper
bound of $4.5 \unit{K}$ is exceeded by a third of the \acs{CMIP}6 models, with
many models showing \ac{ECS} values above $5 \unit{K}$. For \ac{TCR}, the model
range of \acs{CMIP}6 is $\rangeunit{1.3}{3.0}{K}$, which also exceeds the
\acs{CMIP}5 range of $\rangeunit{1.1}{2.5}{K}$ and the assessed range from
\acs{AR}5 of $\rangeunit{1.0}{2.5}{K}$. One possible reason for the increased
climate sensitivity in many \acs{CMIP}6 models is the addition of prognostic
aerosol schemes that include aerosol-cloud interactions, which could have
produced overly large negative radiative forcing. To correctly reproduce the
observed historical temperature trend, a stronger response to \ac{GHG} forcing
is required, resulting in a higher climate sensitivity in the corresponding
models. A further reason for the high \ac{ECS} in \acs{CMIP}6 is a change in
the microphysical representation of mixed-phase clouds over the Southern Ocean
in some models. This change was mainly implemented to reduce a long-standing
cloud phase bias over the Southern Ocean that is present in models from
previous \ac{CMIP} generations. Moreover, the \acs{CMIP}6 \ac{MMM} shows an
improved simulation of the shortwave \ac{CRE} when compared to observations.
However, the change in the representation of mixed-phase clouds also
substantially reduces the strong negative shortwave cloud feedback over the
Southern Ocean that is present in previous \ac{CMIP} generations and that
results from a cloud phase change from ice clouds in the present-day to liquid
clouds in the future. In the affected \acs{CMIP}6 models, this cloud phase
change due to warming is reduced since these models simulate less cloud ice
over the present-day Southern Ocean than their predecessor versions with no
supercooled cloud liquid.
